---
title: "p8105_hw2_fwt2107"
author: "Felix Tran"
date: "September 29, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Problem 1

## Reading and cleaning the dataset

First, we read in the dataset, clean the variable names, select which 
variables to keep, and recode entry from a character variable (YES/NO) to a
logical variable (T/F)
```{r}
mta_df <- read_csv('./Data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv') %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, 
         route1:route11, entry, vending, entrance_type, ada) %>% 
  mutate(entry = ifelse(entry == 'YES', TRUE, FALSE))

mta_df
View(mta_df)
```

This dataset contains information about the entrances for NYC's subway 
stations. Variables include the route name, station name, station coordinates,
routes serviced at the station, whether or not the entrance allows entry into
the station, whether or not the station includes a ticket vending machine,
the type of entrance (e.g. stairs, door, etc.), and whether or not the 
entrance is ADA-compliant. 

The resulting dataset is 1868 rows by 19 columns. 

These data are not tidy because the route variable(s) contain 2 pieces of
information: how many routes are serviced at the particular station, and which
routes are serviced. Someone could tell visually how many routes and which 
routes are serviced at a particular station, but determining that information
in R is not as straightforward given how the data are currently formmated.
These data could be tidied by creating a variable for the number of routes 
serviced at each station and creating variables for each possible route 
(e.g. 'A', '1', 'R', etc.) to indicate if the routes are serviced at each 
particular station. 
