---
title: "p8105_hw2_fwt2107"
author: "Felix Tran"
date: "September 29, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggridges)
```

# Problem 1

## Reading and cleaning the dataset

First, we read in the dataset, clean the variable names, select which 
variables to keep, and recode entry from a character variable (YES/NO) to a
logical variable (T/F)
```{r}
mta_df <- read_csv('./Data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv') %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, 
         route1:route11, entry, vending, entrance_type, ada) %>% 
  mutate(entry = ifelse(entry == 'YES', TRUE, FALSE))

mta_df
View(mta_df)
```

## Describing the data

This dataset contains information about the entrances for NYC's subway 
stations. Variables include the route name, station name, station coordinates,
routes serviced at the station, whether or not the entrance allows entry into
the station, whether or not the station includes a ticket vending machine,
the type of entrance (e.g. stairs, door, etc.), and whether or not the 
entrance is ADA-compliant. 

The resulting dataset is 1868 rows by 19 columns. 

These data are not tidy because each row contain 2 pieces of information: 
how many routes are serviced at the particular station, and which routes are 
serviced. Someone could tell visually how many routes and which routes are 
serviced at a particular station, but determining that information in R is 
not as straightforward given how the data are currently formmated. The data 
could be tidied by making each row one route. The entry variable could also 
be reformatted from character to logical for ease of use in analysis.

## Answering questions

Only keeping distinct values of line and station_name gives us a dataframe 
with 465 rows. There are 465 distinct stations.
```{r}
distinct(mta_df, line, station_name, .keep_all = T)
```

Keeping distinct values of line and station_name plus filtering out stations
which weren't ADA-compliant gives us a dataframe with 84 rows. 84 stations 
are ADA-compliant.
```{r}
distinct(mta_df, line, station_name, .keep_all = T) %>% 
  filter(ada == T)
```

Filtering out observations for entrances/exits without vending gives us a 
dataset with 183 rows. There are 183 entrances/exits without vending.
```{r}
filter(mta_df, vending == 'NO')
```

Filtering out entrances/exits without vending and no entry gives us a dataset
with 69 rows. There are 69 entrances/exits without vending which allow 
entrance. Therefore, 69/183 or about 37.7% of station entrances/exits without 
vending allow entrance.
```{r}
filter(mta_df, vending == 'NO', entry == T)
```

First, we gather the data so that each row corresponds to 1 route. Then we only
keep the observations for the A route. Finally we keep distinct values of line 
and station_name to give us a dataframe with 60 rows. 60 distinct stations 
service the A. 
```{r}
gather(mta_df, key = route_num, value = route_name, route1:route11) %>% 
  filter(route_name == 'A') %>% 
  distinct(line, station_name, .keep_all = T)
```

Copying over our most recent pipeline and then filtering out stations which
are not ADA-compliant gives us a dataframe with 17 rows. 17 distinct stations
which service the A are ADA-compliant.
```{r}
gather(mta_df, key = route_num, value = route_name, route1:route11) %>% 
  filter(route_name == 'A') %>% 
  distinct(line, station_name, .keep_all = T) %>% 
  filter(ada == T)
```

# Problem 2

## Mr. Trash Wheel
First we read in the data. Then we clean the variable names, filter out rows
which are not for specific dumpsters, and change the sports_ball variable into
integer. 
```{r}
trashwheel_df <- 
  readxl::read_excel('./Data/HealthyHarborWaterWheelTotals2017-9-26.xlsx', 
                     range = "A2:N258") %>% 
  janitor::clean_names() %>%
  filter(!is.na(dumpster)) %>% 
  mutate(sports_balls = as.integer(sports_balls))
  
trashwheel_df
```

## Precipitation 2016 and 2017

First we read in the precipitation 2016 and 2017 data, and then we clean the
variable names, add a year variable, and filter out rows with no rain data.
```{r}
precip2016_df <- 
  readxl::read_excel('./Data/HealthyHarborWaterWheelTotals2017-9-26.xlsx',
                     sheet = '2016 Precipitation', range = 'A2:B14') %>% 
  janitor::clean_names() %>% 
  mutate(year = 2016)
precip2016_df

precip2017_df <-
  readxl::read_excel('./Data/HealthyHarborWaterWheelTotals2017-9-26.xlsx',
                     sheet = '2017 Precipitation', range = 'A2:B14') %>% 
  janitor::clean_names() %>% 
  mutate(year = 2017) %>% 
  filter(!is.na(total))
```

We combine the precipitation datasets and convert month to a character
variable.
```{r}
precip_combined_df <- full_join(precip2016_df, precip2017_df) %>% 
  mutate(month = month.name[month])
precip_combined_df
```

## Data descriptions
```{r include = F}
# For Mr. Trash Wheel, filter out observations not from 2016 to make it easier
# to use inline code to describe attributes of the 2016 cleanup.

# For precipitation, filter out observations not from 7 to make it easier
# to use inline code to describe precipitation in 2017.
trashwheel_2016_df <- 
  filter(trashwheel_df, year == '2016') %>% 
  arrange(sports_balls)

precip_2017_tidied <- filter(precip_combined_df, year == 2017)
```

The Mr. Trash Wheel dataset contains `r nrow(trashwheel_df)` observations. 
It contains variables such as the dumpster number, date, and counts of 
particular items in each dumpster such as sports balls, plastic bags, and
cigarette butts.

The combined precipitation 2016-2017 dataset contains 
`r nrow(precip_combined_df)` observations. It contains variables such as the 
month, year, and amount of precipitation in inches per month. 

Total precipitation in 2017 was `r sum(precip_2017_tidied$total)` inches.

The median number of sports balls in a dumpster in 2016 was 
`r median(trashwheel_2016_df$sports_balls)`.

# Problem 3
Load BRFSS data. 
```{r}
library(p8105.datasets)
brfss_df <- brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  select(-class, -topic,  -question, -sample_size, 
         -c(confidence_limit_low:geo_location))


brfss_df
```

